Numerical simulation of the Cordilleran ice sheet through the last glacial cycle

### 2014-06-10 cordillera-cycle.tex
Add discussion of glaciated areas.

These results makes us believe that a palaeo-temperature proxy record located
in the direct vicinity of the former Cordilleran ice sheet would be of great
help to understand its glacial history. Because, to our knowledge, no such
record yet exists, we focus the rest of our analysis on simulations forced by
temperature records from one Greenland (GRIP) and one Antarctic (EPICA) ice
cores.

@@

During most of the glacial cycle, ice cover is restricted to high mountain
areas. This is particularly true when using the Greenland-based forcing
(Fig~\ref{fig:duration}). Most of the glacial maxima margin has a transient
character, showing that the ice sheet is far out of balance with climate. There
persists a central ice dome over the Skeena mountains over the entire glacial
period. This ice dome serves as a nucleation centre at the initiation of the
advance phase preceding the last glacial maximum. It is here because of more
widespread precipitation over the Skeena mountains than across other East-West
transects towards south or north of this region. We interpret this as an
orographic precipitation effect resulting from a "window" through the sharp
range of the Coast Mountains at this latitude (Fig?).

>>>>

These results makes us believe that a palaeo-temperature proxy record located
in the direct vicinity of the former Cordilleran ice sheet would be of great
help to understand its glacial history. Because, to our knowledge, no such
record yet exists, we focus the rest of our analysis on simulations forced by
temperature records from a Greenland (GRIP) and an Antarctic (EPICA) ice cores.

@@

Although snapshots from the model output (Fig.~\ref{fig:duration}) give a
crisp picture of modelled ice dynamics at specific time frames, they often
can not be directly compared to geomorphological evidence, which is by nature
time-transgressive. In this section, we use the 6-km horizontal-resolution
model output from simulations driven by the GRIP and EPICA records to produce
a series of time-integrated plots containing information aggregated over the
whole simulation period.

During most of the glacial cycle, modelled ice cover is restricted to disjoint
ice caps centred on major mountain ranges of the North American Cordillera
(Fig.~\ref{fig:duration}). A~contiguous ice cover spanning from the Alaska
Range to the Rocky Mountains exists for about 28\,\unit{kyr} in the GRIP
simulation and for about 29\,\unit{kyr} in the EPICA simulation. These values
represent cumulative ice cover over the two glaciations during MIS~2 and MIS~4,
while a dynamically consistent Cordilleran ice sheet forms for even shorter
time spans.

With exception of the oceanic front and the northern foothills of the Alaska
Range, most of the maximal extent of the ice sheet is associated with short
durations of ice cover, generally a few thousand years, distributed over one
(GRIP) or two (EPICA) glaciations. This shows that, in our simulations, the
most extensive stages of the Cordilleran ice sheet are both short-lived and
out of balance with contemporary climate states, a result which confirms
inferences from earlier geological studies \aref.

However, our simulations depict, independently of the choice of record used for
palaeo-temperature forcing, a central ice cap persisting throughout the entire
last glacial cycle on the Skeena mountains (Fig.~\ref{fig:snapshots} and
\ref{fig:duration}). In particular, this ice cap persists between the MIS~4
and MIS~2 glaciations, providing a nucleation centre for the onset of glacial
readvance towards the last glacial maximum.

During time periods when a fully-grown Cordilleran ice sheet is absent from our
simulation, the presence of the Skeena ice cap can be explained by more
widespread winter precipitation there than in other parts of the modelling
domain (Figs.~\ref{fig:atm}). Along most of the north-western North American
shore, coastal mountain ranges provide a sharp orographic obstacle to westerly
winds, capturing atmospheric moisture in the form of precipitation, and leaving
interior lowlands arid. However, near the centre of our modelling domain, the
topographic barrier formed by the Coast Mountains is less pronounced than
elsewhere, allowing westerly winds to carry moisture further inland, until it
is progressively captured by the broad and mildly elevated group of the Skeena
Mountains.

In addition, the erosional landscape of the Skeena mountain bears one of the
strongest glacial imprint (?) found within the area formerly covered by the
Cordilleran ice sheet \aref. Here we suggest that persistent ice cover
associated with basal temperatures at the ice melting point
(Fig.~\ref{fig:warmbase}) may partly explain this spectacular landscape.
